---
title: "Big Grid block-level analyses"
author: "Andrew Crosby"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = normalizePath(".."))  #Sets the working directory to the project folder
```


```{r, warning=FALSE, message=FALSE, echo=FALSE}

rm(list=ls(all = TRUE))

library(reshape2)
library(abind)
library(tidyr)
library(jagsUI)
library(boot)
library(coda)
library(knitr)
library(loo)
library(dplyr)

```


The difference between this model and previously-developed mutli-scale occupancy models [@Mordecai2011; @Crosby2018] is that the quantity of interest is not presence/absence at the larger-scale unit level, but rather the proportion of area occupied within the unit as a function of unit-level variables (i.e., human footprint within the unit). Thus, probability of use at the sub-unit level is constrained by unit level human footprint as well as conditional on unit level occupancy. 

In the most basic model, we estimate the proportion of area occupied as a mean probability of occupancy among all sites within a block, which is a function of human footprint + error. We accounted for heterogeneity in occupancy probability by modeling probability of detection as a function of dominant habitat and survey-level covariates julian day and time from sunrise. 



```{r, warning=FALSE, message=FALSE}

load("0_data/processed/big-grid-point-data.Rdata")
load("0_data/processed/big-grid-veghf-blocks.Rdata")



# Get the dominant landcover type within 150m of each point 
d <- veghf_150
colnames(d)[19] <- "prop.hf"
lc <- d[, c("prop.decid", "prop.mixed", "prop.pine", "prop.spruce", "prop.Bspr", "prop.larch", "prop.low", "prop.open")]
is.decid <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 1, 1, 0), 0)))
is.mixed <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 2, 1, 0), 0)))
is.pine <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 3, 1, 0), 0)))
is.spruce <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 4, 1, 0), 0)))
is.Bspr <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 5, 1, 0), 0)))
is.larch <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 6, 1, 0), 0)))
is.low <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 7, 1, 0), 0)))
is.open <- unlist(sapply(1:nrow(lc), function(x) ifelse(max(lc[x, ]) > 0, ifelse(which(lc[x, ] == max(lc[x, ])) == 8, 1, 0), 0)))


veghf150_dom <- data.frame(d, is.decid, is.pine, is.mixed, is.spruce, is.Bspr, is.larch, is.low, is.open)

dom.list <- c("is.decid", "is.pine", "is.mixed", "is.spruce", "is.Bspr", "is.larch", "is.low", "is.open")

# Create the dominance weighting
for(i in 1:length(dom.list)){
  t <- dom.list[i]
  v <- which(colnames(veghf150_dom) == paste0("prop.", substr(dom.list[i], 4, nchar(dom.list[i]))))
  assign(paste0("dom.", substr(dom.list[i], 4, nchar(dom.list[i]))), pmax(0, pmin(1, 2*veghf150_dom[, v] - 0.5)))
}


veghf150_dom <- data.frame(veghf150_dom, dom.decid, dom.pine, dom.mixed, dom.spruce, dom.Bspr, dom.larch, dom.low, dom.open)

# Expand the veg data to add the quadratic and interaction terms
veghf_expand <- data.frame(age = veghf150_dom$wt_mean_age, age2 = veghf150_dom$wt_mean_age^2, veghf150_dom[, c(28:35)], total.hf = veghf_300$total.hf)


# Standardize the age variables 
veghf_dom <- veghf_expand %>% mutate_at(c("age", "age2", "total.hf"), ~(scale(.) %>% as.vector))
rownames(veghf_dom) <- rownames(veghf_expand)

all.equal(rownames(veghf_dom), bg_nsurv_veg$StationKey, bg_ss_veg$StationKey)

```


Run the basic block model where psi is constant based on hf and error, and p varies according to dominant habitat type, jday, and tssr
```{r}


  spp_code <- "OVEN"

  spp_point <- bg_pa[, c(1:2, which(colnames(bg_pa) == spp_code))] 
  
  spp_pt <- spp_point %>% spread(survey, spp_code) 
  spp_pt <- spp_pt[match(bg_nsurv_veg$StationKey, spp_pt$StationKey), 1:5] 
  
  cat("Rows match?\n")
  print(all.equal(bg_nsurv_veg$StationKey, spp_pt$StationKey, rownames(veghf))) 
  
  hab <- veghf150_dom[, c("is.decid", "is.pine", "is.mixed", "is.spruce", "is.Bspr", "is.larch", "is.low", "is.open")]
  
  det <- det.var_veg
  
  all.equal(rownames(det.var_veg), rownames(veghf150_dom))
  d <- list()
  for(i in 1:dim(det.var_veg)[2]){
    b <- cbind(det.var_veg[, i, "int"], hab, det.var_veg[, i, c(2:3)])
    colnames(b) <- c("int", colnames(b[-1]))
    d[[i]] <- b
  }
  
  
  
  det.hab <- abind::abind(d, along = 3)
  det.hab <- aperm(det.hab, c(1, 3, 2))
  
  
  all.equal(rownames(det.var_veg), rownames(veghf150_dom))

```


Run the test model for oven bird in the 2x2 blocks

```{r}

blockhf <- grids_veghf[[2]][, c("blockID", "total.hf")]

block_cov <- data.frame(int = rep(1, nrow(blockhf)), blockhf[, "total.hf"])
rownames(block_cov) <- blockhf$blockID

point_cov <- data.frame(int = rep(1, nrow(hab)), hab)

all.equal(spp_pt$StationKey, rownames(det), rownames(point_cov))

set.seed(1234)
test <- sort(sample(1:nrow(blockhf), 50, replace = FALSE))

block.test <- blockhf[test, ]
block_cov.test <- block_cov[test, ]
rownames(block_cov.test) <- rownames(block_cov)[test]

pt.test <- point.block[[2]][which(point.block[[2]]$block %in% block.test$blockID), ]

spp_pt.test <- spp_pt[match(pt.test$ss, spp_pt$StationKey), ]
det.test <- det[match(pt.test$ss, rownames(det)), , 1:3]
point_cov.test <- point_cov[match(pt.test$ss, rownames(point_cov)), , ]
nsurv.test <- bg_nsurv_veg[match(pt.test$ss, bg_nsurv_veg$StationKey), ]

all.equal(spp_pt.test$StationKey, rownames(det.test), rownames(hab.test), pt.test$ss, nsurv.test$StationKey)


ni <- 1000; nt <- 1; nb <- 500; nc <- 3 
    
params_1 <- c("beta.block", "beta.point", "beta.p", "sd.grid", "l.score")#, "lprob.y") 
zst <- apply(spp_pt.test[, -1], 1, max, na.rm = T)  
inits.psi <- function() {list(z = zst, beta.block = rnorm(ncol(block_cov.test)), beta.point = rnorm(ncol(point_cov.test)), beta.p = rnorm(dim(det.test)[3]))}    
    
data <- list(n.site = nrow(spp_pt.test), y = as.matrix(spp_pt.test[, -1]), cov.block = block_cov.test, cov.point = point_cov.test, cov.p = det.test, n.beta.block = ncol(block_cov.test), n.beta.point = ncol(point_cov.test), n.beta.p = dim(det.test)[3], n.surv = nsurv.test$nsurv, n.block = nrow(block_cov.test), block = as.numeric(as.factor(rownames(block_cov.test))), point.block = as.numeric(as.factor(pt.test$block)), indicator = zst, n.grid = nlevels(as.factor(pt.test$grid)), point.grid = as.numeric(as.factor(pt.test$grid)))

  system.time({
    out_1 <- jags(data = data, inits = inits.psi, parameters.to.save =  params_1, model.file =  "1_scripts/model_scripts/block_occupancy_1.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE)
    })
  
waic_1 <- waic(out_1$sims.list$lprob.y) 

params_2 <- c("beta.block", "alpha", "beta1", "beta.point", "beta.p", "l.score", "lprob.y") 
inits_2 <- function() {list(z = zst, beta.block = rnorm(ncol(block_cov.test)), beta.point = rnorm(ncol(point_cov.test) - 1), beta.p = rnorm(dim(det.test)[3]))}    

data_2 <- list(n.site = nrow(spp_pt.test), y = as.matrix(spp_pt.test[, -1]), cov.block = block_cov.test, cov.point = point_cov.test[, -1], cov.p = det.test, n.beta.block = ncol(block_cov.test), n.beta.point = ncol(point_cov.test) - 1, n.beta.p = dim(det.test)[3], n.surv = nsurv.test$nsurv, n.block = nrow(block_cov.test), block = as.numeric(as.factor(rownames(block_cov.test))), point.block = as.numeric(as.factor(pt.test$block)), indicator = zst)


  system.time({
    out_2 <- jags(data = data_2, inits = inits_3, parameters.to.save =  params_2, model.file =  "1_scripts/model_scripts/block_occupancy_2.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE)
    })
  
waic_2 <- waic(out_2$sims.list$lprob.y) 

  
inits_3 <- function() {list(z = zst)}    

params_3 <- c("beta.block", "beta.p", "l.score", "lprob.y") 

  system.time({
    out_3 <- jags(data = data, inits = inits_3, parameters.to.save =  params_3, model.file =  "1_scripts/model_scripts/block_occupancy_3.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE)
    })
  
waic_3 <- waic(out_3$sims.list$lprob.y) 

  

```

Run the cross-validation model
```{r}
nobs.block <- nrow(block.test)
k = 5
nk <- 1/k
leavout <- sample(rep(1:(nobs.block/(nobs.block*nk)), ceiling(nobs.block*nk))[1:nobs.block])

sapply(1:length(unique(leavout)), function(x) length(which(leavout == x)))
j <- 1

test <- which(leavout == j)
block.train <- block.test[-test, ]
block.cv <- rbind(block.train, block.test[test, ])

block_cov.cv <- rbind(block_cov.test[-test, ], block_cov.test[test, ])

pt.train <- pt.test[which(pt.test$block %in% block.train$blockID), ]
point.test <- pt.test[which(!(pt.test$block %in% block.train$blockID)), ]
pt.cv <- rbind(pt.train, point.test)

spp_pt.cv <- spp_pt[match(pt.cv$ss, spp_pt$StationKey), ]
det.cv <- det[match(pt.cv$ss, rownames(det)), , 1:3]
point_cov.cv <- point_cov[match(pt.cv$ss, rownames(point_cov)), , ]
nsurv.cv <- bg_nsurv_veg[match(pt.cv$ss, bg_nsurv_veg$StationKey), ]

all.equal(spp_pt.cv$StationKey, rownames(det.cv), rownames(hab.cv), pt.cv$ss, nsurv.cv$StationKey)


ni <- 1000; nt <- 1; nb <- 500; nc <- 3 
    
params_1 <- c("beta.block", "beta.point", "beta.p", "l.score") 
zst <- apply(spp_pt.cv[, -1], 1, max, na.rm = T)  
inits_1 <- function() {list(z = zst[1:nrow(pt.train)], beta.block = rnorm(ncol(block_cov.cv)), beta.point = rnorm(ncol(point_cov.cv)), beta.p = rnorm(dim(det.cv)[3]))}    
    
data <- list(ntrain.block = nrow(block.train), ntrain.site = nrow(pt.train), ntest.block = length(test), ntest.site = nrow(pt.cv) - nrow(pt.train), y = as.matrix(spp_pt.cv[, -1]), cov.block = block_cov.cv, cov.point = point_cov.cv, cov.p = det.cv, n.beta.block = ncol(block_cov.cv), n.beta.point = ncol(point_cov.cv), n.beta.p = dim(det.cv)[3], n.surv = nsurv.cv$nsurv, n.block = nrow(block_cov.cv), block = as.numeric(as.factor(rownames(block_cov.cv))), point.block.train = as.numeric(as.factor(pt.train$block)), point.block.test = as.numeric(as.factor(point.test$block)), indicator = zst)

  system.time({
    out_1.cv <- jags(data = data, inits = inits_1, parameters.to.save =  params_1, model.file =  "1_scripts/model_scripts/block_occupancy_1_cv.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE)
    })

```

Turn it into a function using the Bayesian Lasso
```{r}

nlam <- 20

log.lambda=seq(-5, 5, length = nlam)
lambda <- exp(log.lambda)

plot(1:nlam,1/lambda,type="l") 


cv.loop_lasso <- function(blocks, blocks_cov, points, points_cov, spp, det, nsurv, k, ni, nt, nb, nc, lambda){
  nobs.block <- nrow(block.test)
  nk <- 1/k
  leavout <- sample(rep(1:(nobs.block/(nobs.block*nk)), ceiling(nobs.block*nk))[1:nobs.block])
  
  score <- sapply(1:k, function(j){
    test <- which(leavout == j)
    b.t <- blocks[-test, ]
    b.cv <- rbind(b.t, blocks[test, ])
    
    bcov.cv <- rbind(block_cov[-test, ], block_cov[test, ])
    
    pts.t <- points[which(points$block %in% b.t$blockID), ]
    pts.test <- points[which(!(points$block %in% b.t$blockID)), ]
    pts.cv <- rbind(pts.t, pts.test)
    
    spp_pt.cv <- spp[match(pts.cv$ss, spp$StationKey), ]
    det.cv <- det[match(pts.cv$ss, rownames(det)), , 1:3]
    points_cov.cv <- points_cov[match(pts.cv$ss, rownames(points_cov)), , ]
    nsurv.cv <- nsurv[match(pts.cv$ss, nsurv$StationKey), ]
    
    all.equal(spp_pt.cv$StationKey, rownames(det.cv), rownames(points_cov.cv), pts.cv$ss, nsurv.cv$StationKey)
    
    params <- c("beta.block", "beta.point", "beta.p", "l.score") 
    zst <- apply(spp_pt.cv[, -1], 1, max, na.rm = T)  
    inits <- function() {list(z = zst[1:nrow(pts.t)], beta.block = rnorm(ncol(bcov.cv)), beta.point = rnorm(ncol(points_cov.cv)), beta.p = rnorm(dim(det.cv)[3]))}    
    
    data <- list(ntrain.block = nrow(b.t), ntrain.site = nrow(pts.t), ntest.block = length(test), 
                 ntest.site = nrow(pts.cv) - nrow(pts.t), y = as.matrix(spp_pt.cv[, -1]), cov.block = bcov.cv, 
                 cov.point = points_cov.cv, cov.p = det.cv, n.beta.block = ncol(bcov.cv), 
                 n.beta.point = ncol(points_cov.cv), n.beta.p = dim(det.cv)[3], n.surv = nsurv.cv$nsurv, 
                 n.block = nrow(bcov.cv), block = as.numeric(as.factor(rownames(bcov.cv))), 
                 point.block.train = as.numeric(as.factor(pts.t$block)), 
                 point.block.test = as.numeric(as.factor(pts.test$block)), indicator = zst, lambda = lambda)
    
    system.time({
      out <- jags(data = data, inits = inits, parameters.to.save =  params_1, 
                  model.file =  "1_scripts/model_scripts/block_occupancy_1_cv_lasso.txt", 
                  n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE)
    })
    return(out$mean$l.score)
  })
  return(score)
}

system.time({
  lasso.1 <- mean(cv.loop_lasso(blocks = block.test, blocks_cov = block_cov.test, points = pt.test, points_cov = point_cov, spp = spp_pt, det = det, nsurv = bg_nsurv_veg, k = 5, ni = 1000, nt = 1, nb = 500, nc = 3, lambda = lambda[1]))
})

system.time({
  lasso.cv <- sapply(1:length(lambda), function(i) 
                mean(cv.loop_lasso(blocks = block.test, blocks_cov = block_cov.test, points = pt.test, points_cov = point_cov, spp = spp_pt, det = det, nsurv = bg_nsurv_veg, k = 5, ni = 1000, nt = 1, nb = 500, nc = 3, lambda = lambda[i]))
                )
})

l1 <- which(lasso.cv == min(lasso.cv))

lasso.best <- function(blocks, blocks_cov, points, points_cov, spp, det, nsurv, ni, nt, nb, nc, lambda){
  params_1 <- c("beta.block", "beta.point", "beta.p", "l.score", "lprob.y") 
  zst <- apply(spp[, -1], 1, max, na.rm = T)  
  inits.psi <- function() {list(z = zst, beta.block = rnorm(ncol(blocks_cov)), beta.point = rnorm(ncol(points_cov)), beta.p = rnorm(dim(det)[3]))}    
  
  data <- list(n.site = nrow(spp), y = as.matrix(spp[, -1]), cov.block = blocks_cov, cov.point = points_cov, cov.p = det, n.beta.block = ncol(blocks_cov), n.beta.point = ncol(points_cov), n.beta.p = dim(det)[3], n.surv = nsurv$nsurv, n.block = nrow(blocks_cov), block = as.numeric(as.factor(rownames(blocks_cov))), point.block = as.numeric(as.factor(points$block)), indicator = zst, lambda = lambda, n.grid = nlevels(as.factor(points$grid)), point.grid = as.numeric(as.factor(points$grid)))

  system.time({
    out_1 <- jags(data = data, inits = inits.psi, parameters.to.save =  params_1, model.file =  "1_scripts/model_scripts/block_occupancy_1_lasso.txt", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, parallel = TRUE)
    })
  
  return(out_1)
}

system.time({
  lasso.model <- lasso.best(blocks = block.test, blocks_cov = block_cov.test, points = pt.test, points_cov = point_cov, spp = spp_pt.test, det = det.test, nsurv = nsurv.test, ni = 1000, nt = 1, nb = 500, nc = 3, lambda = lambda[l1])
})


lasso.waic <- waic(lasso.model$sims.list$lprob.y)


```


Now test it with all the variables
```{r}

blockhf <- grids_veghf[[2]] %>% select(blockID, prop.seismic, prop.widelin, prop.wells, prop.industry)


block_cov <- data.frame(int = rep(1, nrow(blockhf)), blockhf %>% select(-blockID))
rownames(block_cov) <- blockhf$blockID

point_cov <- data.frame(int = rep(1, nrow(hab)), hab)

all.equal(spp_pt$StationKey, rownames(det), rownames(point_cov))

set.seed(1234)
test <- sort(sample(1:nrow(blockhf), 100, replace = FALSE))

block.test <- blockhf[test, ]
block_cov.test <- block_cov[test, ]
rownames(block_cov.test) <- rownames(block_cov)[test]

pt.test <- point.block[[2]][which(point.block[[2]]$block %in% block.test$blockID), ]

spp_pt.test <- spp_pt[match(pt.test$ss, spp_pt$StationKey), ]
det.test <- det[match(pt.test$ss, rownames(det)), , 1:3]
point_cov.test <- point_cov[match(pt.test$ss, rownames(point_cov)), , ]
nsurv.test <- bg_nsurv_veg[match(pt.test$ss, bg_nsurv_veg$StationKey), ]

all.equal(spp_pt.test$StationKey, rownames(det.test), rownames(hab.test), pt.test$ss, nsurv.test$StationKey)

system.time({
  lasso.cv <- sapply(1:length(lambda), function(i) 
                mean(cv.loop_lasso(blocks = block.test, blocks_cov = block_cov.test, points = pt.test, points_cov = point_cov, spp = spp_pt, det = det, nsurv = bg_nsurv_veg, k = 5, ni = 1000, nt = 1, nb = 500, nc = 3, lambda = lambda[i]))
                )
})

l1 <- which(lasso.cv == min(lasso.cv))

system.time({
  lasso.model <- lasso.best(blocks = block.test, blocks_cov = block_cov.test, points = pt.test, points_cov = point_cov, spp = spp_pt.test, det = det.test, nsurv = nsurv.test, ni = 1000, nt = 1, nb = 500, nc = 3, lambda = lambda[l1])
})


lasso.waic <- waic(lasso.model$sims.list$lprob.y)

```


## Literature Cited